Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-02 files/Overview.tex
overview v2

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and in turn a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of the chapter, from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. The assumptions and results can be formulated and applied to all the applications.

In this section, the existence of a unique solutions for a general vibration problem is investigated. The main article that is discussed is \cite{VV02}. In this article, the authors define a general vibration problem. The authors state four assumptions that need to be satified. Under these four assumptions, it can be shown that for the general vibration problem there exists a linear operator A, which is an infinitesimal generator of a $C_0$-semigroup of contractions with a domain in the problem space. This operator allows the general vibration problem to be rewritten as a system of first order differential equations $\displaystyle x' = Ax + f$. This differential equations are rewritten into an initial value problem which can be proven using semi-group theory to have a unique solution. The authors then provide the necessary proofs to show that a solution of the initial value problem is also a solution of the general vibration problem. An example of the application of this theory is provided using a cantilever Timoshenko beam.

\subsubsection{Chapter 3}
In this chapter, some theoretical results for the Finite Element Method is discussed. This chapter contains two parts, that covers two results. The first result is the convergence of the Galerkin approximation of a general vibration problem. This result is presented in the article \cite{BV13}. The second result concerns the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. This result is presented in the textbook \cite{SF97}. In the first section, the authors continue with the general vibration problem that is used in chapter 2, as well as the assumptions made in chapter 2. Hence there is a unique solution. The Galerkin Approximation for the general vibration problem is derived and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The goal of \cite{BV13} is to show that the solution for the general vibration problem and the solution for the Galerkin Approximation can be made arbitrarily close by choosing a sufficiently small $h$. The approach of the authors is to calculate the error estimates. The error is split into two parts, a semi-discrete and a fully discrete problem. Two new assumptions are added to the assumptions from chapter 2. Using these assumptions, the authors obtain error estimates for the semi-discrete and fully discrete problems. This result shows that the error can be made arbitrarily small by choosing a sufficiently small $h$.

@@

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this disseration, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models. The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and look at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

@@

\section{500-word summary}
This dissertation is a literature study that investigates the validity of different linear models for applications. Validity in this context refers to how well simplified lower dimensional models compare to more realistic higher dimensional models. This dissertation will investigate the validity of the Timoshenko beam theory and the Reissner-Mindlin plate theory, with focus on the Timoshenko theory. The higher dimensional models are a two- and three-dimensional beam models and a three-dimensional plate model.

In the first chapter, the models are provided. Then a discussion follows of the existence and uniqueness of solutions for a general vibration problem. The article discussed provide necessary four assumptions, and prove there exists an infinitesimal generator of a $C_0$-semigroup of contractions, that rewrites the problem into an initial value problem, with a unique solution. The authors then prove this is also a solution of the general vibration problem. An application of this theory is demonstrated using a cantilever Timoshenko beam.

Thereafter of two theoretical results for the Finite Element Method is discussed, one from an article and one from a textbook. The first result looks at the convergence of the Galerkin Approximation for vibration problems. Using the Finite Element Method, this Galerkin Approximation is rewritten into an initial value problem, that has a unique solution. The authors show that there exists error estimates that approximate the error between the general vibration problem and the initial value problem. The next result is on the convergence of the eigenvalues and eigenvectors of a one-dimensional vibration problem. This section rewrites the textbook's results with updated notation, aiming for greater clarity and understanding.

Next the focus turns to the Timoshenko beam theory. First an article is examined that provides a method to calculate the exact eigenvalues and eigenvectors. The method is then applied to a cantilever and a free-free beam model. Next is a discussion of an empirical study, that compares the natural frequencies of a physical free-free beam to the theoretical results of the Timoshenko beam theory. Finally an extension on a technical report of a cantilever Timoshenko beam with a tip-body and elastic interfaces. This model serves as an example of the application of the Timoshenko theory in a complex model and adds to the article by expanding the numerical results.

The attention is then turned to application where the Finite Element Method is applied to the two and three-dimensional models of this dissertation. The goal of this chapter is to obtain a numerical formula that can be used to calculate the eigenvalues and eigenvectors of the models in preparation for the final chapter where the models are compared.

The final chapter compares the eigenvalues and eigenvectors of the models. The approach is based on an article where a cantilever Timoshenko beam model is compared to a cantilever two-dimensional model. The article is discussed and results replicated. The article is then extended to a comparison of a two-dimensional model to a three-dimensional model. Finally, a cantilever Reissner-Mindlin plate model is compared to a cantilever three-dimensional plate model.

\textcolor{red}{500 Words}

@@

\subsubsection{Comparison of models}
The first chapter introduces the models used in this dissertation. The models are of different dimensions. The first model is a three dimensional elastic body. The model problems for the three-dimensional body is a cantilever and a free-free body. A general case of plane stress is also discussed. Making assumptions on plane stress, results into a two-dimensional model. This model is a special case of plane stress. For the one-dimensional case, the Timoshenko beam theory is considered. Four different models problems are used in this disseration. These model problems are a cantilever, a free-free beam, a pinned-pinned beam and a suspended beam. Lastly a Reissner Mindlin plate theory is considered. The model problems for the plate theory are a cantilever plate and a rigidly clamped plate. For each of the models problems, the boundary conditions are presented and the variational form is derived. The variational form of each problem is mainly used in this dissertation. (ROD THEORY?)

>>>>

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and in turn a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of the chapter, from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. Therefore the assumptions and results can be formulated and applied to all the applications. The article gives four assumptions. Under these assumptions it is shown that the general vibration problem has a unique solution using semi-group theory. Next the theory is applied to the example weak variational problem for the cantilever Timoshenko beam, by showing that the assumptions hold.

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. The idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed, following \cite{CVV18}. Given a general vibration problem, an eigenvalue problem and a differential equation is derived. An additional assumption is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has infinite many solutions, and that the solution of the general vibration problem is an infinite series of the eigenvectors and general Fourier coefficients. Solving the differential equation, shows that these Fourier coeffiencients are periodic and a formal series solution for the general vibration problem (which is a boundary value problem) is obtained. This formal series solution is then shown to also be valid for a general initial value problem. This general case is applicable to all the models in this dissertation.

Being able to express the solution of a model problem as a linear combination of the eigenvalues and eigenfunctions in a formal series solution, is fundamental to this dissertation. This means that the solutions of models will compare well if the eigenvalues and eigenfunctions compare well. The validity of the formal series solution is also a requirement as it ensures that the comparisons remain valid, as long as the models are disturbed in the same way.

\subsubsection{Chapter 3}
In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different results. The first result is the convergence of the Galerkin approximation of a general vibration problem. This result is presented in the article \cite{BV13}. The second result concerns the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. This result is presented in the textbook \cite{SF97}. In the first section, the authors continue with the general vibration problem that is used in chapter 2. The general vibration problem has a solution as shown in chapter 2. The Galerkin Approximation is derived from the general vibration probelm and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The main results of \cite{BV13} shows that this solution of the Galerkin Approximation converges to the solution of the general vibration problem. The approach of the authors is to calculate the error estimates. The error is split into two parts, a semi-discrete and a fully discrete problem. Two assumptions are added to the assumptions from chapter 2. Using these assumptions, the authors obtain error estimates for the semi-discrete and fully discrete problems.

@@

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this dissertation, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models. The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and look at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

@@

\section{500-word summary}
This dissertation is a literature study that investigates the validity of different linear models in application. Validity in this context refers to how well simplified lower dimensional models compare to more realistic higher dimensional models. This dissertation will investigate the validity of the Timoshenko beam theory and the Reissner-Mindlin plate theory. The higher dimensional models are a two- and three-dimensional beam and a three-dimensional plate. In the first chapter, the models are provided.

Next, a discussion of the existence and uniqueness of solutions for a general vibration problem. The article discussed provide four assumptions, and prove there exists an infinitesimal generator of a $C_0$-semigroup of contractions, that rewrites the problem into an initial value problem, with a unique solution. The authors then prove this is also a solution of the general vibration problem. An application of this theory is demonstrated using a cantilever Timoshenko beam.

Thereafter a discussion of two theoretical results for the Finite Element Method. The first result is from an article where the convergence of the Galerkin Approximation is proved. A Galerkin Approximation of the general vibration problem is derived. Using the Finite Element Method, this Galerkin Approximation is rewritten into an initial value problem, that has a unique solution. The authors show that there exists error estimates that approximate the error between the general vibration problem and the initial value problem. The next result is from a textbook on the convergence of the eigenvalues and eigenvectors of a one-dimensional vibration problem. This section rewrites the textbook's results with updated notation, aiming for greater clarity and understanding.

The forth chapter discusses results for the Timoshenko beam theory. First an article is examined that provides a method to calculate the exact eigenvalues and eigenvectors. The method is then applied to a cantilever and a free-free beam model. Next is a discussion of an empirical study, that compares the natural frequencies of a physical free-free beam to the theoretical results of the Timoshenko beam theory. Finally an extension on a technical report of a cantilever Timoshenko beam with a tip-body and elastic interfaces. This model serves as an example of the application of the Timoshenko theory in a complex model and adds to the article by expanding the numerical results.

In the fifth chapter, the Finite Element Method is applied to the two and three-dimensional models of this dissertation. The goal of this chapter is to obtain a numerical formula that can be used to calculate the eigenvalues and eigenvectors of the models in preparation for the final chapter where the models are compared.

The last chapter compares the eigenvalues and eigenvectors of the models. The approach is based on an article where a cantilever Timoshenko beam model is compared to a cantilever two-dimensional model. The article is discussed and results replicated. The article is then extended to a comparison of a two-dimensional model to a three-dimensional model. Finally, a cantilever Reissner-Mindlin plate model is compared to a cantilever three-dimensional plate model.

\textcolor{red}{493 Words}

@@

\subsubsection{Comparison of models}
The first chapter introduces the models used in this dissertation. The models are of different dimensions. The first model is a three dimensional elastic body. The model problems for the three-dimensional body is a cantilever and a free-free body. A general case of plane stress is also discussed. Making assumptions on plane stress, results into a two-dimensional model. This model is a special case of plane stress. For the one-dimensional case, the Timoshenko beam theory is considered. Four different models problems are used in this dissertation. These model problems are a cantilever, a free-free beam, a pinned-pinned beam and a suspended beam. Lastly a Reissner Mindlin plate theory is considered. The model problems for the plate theory are a cantilever plate and a rigidly clamped plate. For each of the models problems, the boundary conditions are presented and the variational form is derived. The variational form of each problem is mainly used in this dissertation. (ROD THEORY?)
